---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
  
  
```{r, echo=FALSE, message=FALSE, results=FALSE}
library(sqldf)
library(dplyr)
library(knitr)
sqldf <- function(...) knitr::kable(sqldf::sqldf(...))
#source("../dataprep.R")
#source("../ejscreen.R")
load("/tmp/scratch")
```  
  
# Descriptives to re-run with final data

## Number of carceral facilities with unique HIFLD IDs 

```{r}
sqldf("select count(distinct HIFLD_FACILITYID) from facilities")
```



## Number of facilities with unique HIFLD IDs that have corresponding FRS IDs; number of unique FRS IDs  

```{r}
sqldf("
select count(distinct HIFLD_FACILITYID), count(distinct FRS_ID)
from facilities
join frs_hifld using(HIFLD_FACILITYID)       
")
```




## Number of facilities with capacity >25 without FRS IDs; number of facilities with capacity >25 with FRS IDs

```{r}
sqldf("
select frs_hifld.FRS_ID IS NOT NULL as has_FRS,
       count(distinct facilities.HIFLD_FACILITYID),
        count(distinct frs_hifld.FRS_ID)
from facilities
left join frs_hifld on (facilities.HIFLD_FACILITYID = frs_hifld.HIFLD_FACILITYID)
where HIFLD_CAPACITY_2020 >= 25
")
```



## Number of facilities with FRS IDs that have an SDWIS flag, and the corresponding number of unique HIFLD IDs for those facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select PWSID IS NOT NULL as has_pws, count(distinct HIFLD_FACILITYID), count(distinct FRS_ID), count(1)
  from facs
  group by 1
  
  ")
```

## Number of facilities in the Prisons Boundaries dataset (Nick’s dataset) that were missing information about security level or facility type or both. 

```{r}
sqldf("
select SECURELVL IS NOT NULL as has_securelvl, TYPE IS NOT NULL as has_type,
       count(distinct facilities.HIFLD_FACILITYID)
from facilities
group by 1,2
")
```



## Total number of SDWA violations across all facilities in our sample 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select count(distinct VIOLATION_ID)
  from facs
  join violations using(PWSID)
")
```



## number of those violations categorized by type
The number of those violations categorized as monitoring and reporting violations 
The number of those violations categorized as health-based violations (including acute health-based) 
The number of those violations categorized as public notifications violations. 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select ACUTE_HEALTH_BASED = 'Y' or HEALTH_BASED = 'Y' as health,
  MONITORING_REPORTING = 'Y' as monitoring,
  PUBLIC_NOTIF_OTHER = 'Y' as notif,
    count(distinct VIOLATION_ID), count(distinct HIFLD_FACILITYID)
  from facs
  join violations using(PWSID)
  group by 1,2,3
")
```





## The state with the most M&R violations; the number of M&R violations in that state 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select state, count(distinct VIOLATION_ID), count(distinct HIFLD_FACILITYID)
  from facs
  left join violations on(facs.PWSID = violations.PWSID)
  where   MONITORING_REPORTING = 'Y' or MONITORING_REPORTING IS NULL
  group by 1
  order by 2
")
```





## The state with the most health-based violations; the number of health violations in that state 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select state, count(distinct VIOLATION_ID), count(distinct HIFLD_FACILITYID)
  from facs
  left join violations using(PWSID)
  where   ACUTE_HEALTH_BASED is NULL or ACUTE_HEALTH_BASED = 'Y' or HEALTH_BASED = 'Y'
  group by 1
  order by 2
")
```






## The five states with the most violations and the number of those violations; the states with no recorded violations 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select state, count(distinct VIOLATION_ID), count(distinct HIFLD_FACILITYID)
  from facs
  left join violations using(PWSID)
  group by 1
  order by 2
")
```




## Total number of enforcement actions for carceral facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select count(distinct ENFORCEMENT_ID), count(distinct HIFLD_FACILITYID)
  from facs
  join enforcements using(PWSID)
")
```


## Number administered by the state; number administered by the EPA 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select agency, count(distinct ENFORCEMENT_ID), count(distinct HIFLD_FACILITYID)
  from facs
  join enforcements using(PWSID)
  group by 1
  order by 2
")
```


## Total number of formal and number of informal enforcement actions 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select ENFORCEMENT_CATEGORY, count(distinct ENFORCEMENT_ID), count(distinct HIFLD_FACILITYID)
  from facs
  join enforcements using(PWSID)
  group by 1
  order by 2
")
```


## Number of enforcement actions public notifications requested / recvd 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select DESCRIPTION, count(distinct ENFORCEMENT_ID), count(distinct HIFLD_FACILITYID)
  from facs
  join enforcements using(PWSID)
  where lower(DESCRIPTION) LIKE '%notif%'
  group by 1
  order by 2
")
```



## Number of penalties issued 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select ENFORCEMENT_ACTION_TYPE_CODE, count(distinct ENFORCEMENT_ID), count(distinct HIFLD_FACILITYID)
  from facs
  join sdwa_violations_enforcement using(PWSID)
  where ENFORCEMENT_ACTION_TYPE_CODE = 'SFO'
  group by 1
  order by 2
")
```


## Total number of site visits over the 10 years 

```{r}
sqldf("
  with facs as (
    select distinct PWSID
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select 1 as all_visits, count(), count(distinct PWSID)
  from facs
  join visits using(PWSID)
  group by 1
  order by 2
")
```
## Highest number of site visits in a single fiscal year; lowest number of site visits in a single fiscal year. 

```{r}
sqldf("
  with facs as (
    select distinct PWSID
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select FISCAL_YEAR, count(), count(distinct PWSID)
  from facs
  join visits using(PWSID)
  group by 1
  order by 1
")
```

## Absolute numbers and % distribution of different values for ENF_ACTION_CATEGORY for each of the 10 years for carceral facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ),
  freq as (
    select FISCAL_YEAR, ENF_ACTION_CATEGORY, count(1) as n
    from facs
    join sdwa_violations_enforcement using (PWSID)
    join violations using(VIOLATION_ID)
    group by 1,2
    order by 1,2
  )
  select freq.*, round(n * 1.0 / SUM(n) OVER (PARTITION BY FISCAL_YEAR), 4) as perc
  from freq
")
```

## Absolute numbers and % distribution of different values for ENFORCEMENT_ACTION_TYPE_CODE for each of the 10 years for carceral facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ),
  freq as (
    select FISCAL_YEAR, ENFORCEMENT_ACTION_TYPE_CODE, count(1) as n
    from facs
    join sdwa_violations_enforcement using (PWSID)
    join violations using(VIOLATION_ID)
    group by 1,2
    order by 1,2
  )
  select freq.*, round(n * 1.0 / SUM(n) OVER (PARTITION BY FISCAL_YEAR), 4) as perc
  from freq
")
```

## Absolute numbers and % distribution of different values for VIOLATION_STATUS for each of the ten years for carceral facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ),
  freq as (
    select FISCAL_YEAR, VIOLATION_STATUS, count(1) as n
    from facs
    join sdwa_violations_enforcement using (PWSID)
    join violations using(VIOLATION_ID)
    group by 1,2
    order by 1,2
  )
  select freq.*, round(n * 1.0 / SUM(n) OVER (PARTITION BY FISCAL_YEAR), 4) as perc
  from freq
")
```



## Total number of IS_MAJOR_VIOL_IND values Y


```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ),
  freq as (
    select FISCAL_YEAR, IS_MAJOR_VIOL_IND, count(1) as n
    from facs
    join sdwa_violations_enforcement using (PWSID)
    join violations using(VIOLATION_ID)
    group by 1,2
    order by 1,2
  )
  select freq.*, round(n * 1.0 / SUM(n) OVER (PARTITION BY FISCAL_YEAR), 4) as perc
  from freq
")
```


## Absolute numbers and % distribution of different values for VISIT_REASON_CODE and AGENCY_TYPE_CODE for each of the ten years for carceral facilities 

*NB: these columns are only from the other version of visits (SDWA_latest_downloads.zip), they are not present in the one used in previous analysis (SDWA_downloads.zip), and may have different # of rows, coverage, etc. They are groped by visit year, which may be different from fiscal year. etc*



```{r}
#visits2 <- fread(cmd="unzip -p ../data/SDWA_latest_downloads.zip SDWA_SITE_VISITS.csv")
sqldf("
  with facs as (
    select distinct pwsid
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ),
  freq as (
    select substr(visit_date, 7, 10) as VISIT_YEAR, VISIT_REASON_CODE, count(1) as n
    from facs
    join visits2 using (PWSID)
    group by 1,2
    order by 1,2
  )
  select freq.*, round(n * 1.0 / SUM(n) OVER (PARTITION BY VISIT_YEAR), 4) as perc
  from freq
  where VISIT_YEAR >= '2010'
")
```


## Proportion of health-based versus monitoring/reporting violations in carceral versus non-carceral facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select HIFLD_FACILITYID is NOT NULL as carceral, ACUTE_HEALTH_BASED = 'Y' or HEALTH_BASED = 'Y' as health,
  MONITORING_REPORTING = 'Y' as monitoring,
  PUBLIC_NOTIF_OTHER = 'Y' as notif,
    count(distinct VIOLATION_ID), count(distinct HIFLD_FACILITYID)
  from violations 
  left join facs using(PWSID)
  group by 1,2,3,4
")
```


## Ratio of formal to informal enforcement actions in carceral versus non-carceral facilities 


```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select HIFLD_FACILITYID is NOT NULL as carceral, 
  ENFORCEMENT_CATEGORY,
    count(distinct ENFORCEMENT_ID), count(distinct HIFLD_FACILITYID)
  from enforcements 
  left join facs using(PWSID)
  group by 1,2
")
```



## Number of formal enforcement actions at juvenile facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select HIFLD_FACILITYID is NOT NULL as carceral, 
  ENFORCEMENT_CATEGORY,
    count(distinct ENFORCEMENT_ID), count(distinct HIFLD_FACILITYID)
  from enforcements 
  left join facs using(PWSID)
  where SECURELVL = 'JUVENILE'
  group by 1,2
")
```



## Total number of violations in jails (county and local) 


```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select TYPE,
    count(distinct VIOLATION_ID), count(distinct HIFLD_FACILITYID)
  from violations 
   join facs using(PWSID)
  group by 1
")
```



## Total number of violations in ICE facilities 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select ICE_FACILITY,
    count(distinct VIOLATION_ID), count(distinct HIFLD_FACILITYID)
  from violations 
  left join facs using(PWSID)
  group by 1
")
```





## Percentage of facilities that received one or more site visits over the 10 year period 


*NB: They visit PWSs and not facilities*

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  )
  select visits.PWSID IS NOT NULL as visited, count(distinct HIFLD_FACILITYID) 
  from facs 
  left join visits using(PWSID)  
  group by 1
")
```


## Percentage of facilities that received a site visit in each of the 10 years


```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ),
  facyears as (
    select *
    from facs, years
  )
  select FISCAL_YEAR, AVG(visits.PWSID IS NOT NULL) as perc 
  from facyears
  left join visits using(PWSID, FISCAL_YEAR)  
  group by 1
")
```

## The total number of people (minimum estimate) affected by health-based violations in carceral facilities and 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ), caps as (
    select HIFLD_FACILITYID, HIFLD_CAPACITY_2020
    from violations 
    join facs using(PWSID)
    where HEALTH_BASED = 'Y'
        or ACUTE_HEALTH_BASED = 'Y'
    group by 1,2
  )
  select sum(HIFLD_CAPACITY_2020) as health
  from caps
")
```
## the total number of people (minimum estimate) affected by monitoring-based violations in carceral facilities over the last 10 years
```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ), caps as (
    select HIFLD_FACILITYID, HIFLD_CAPACITY_2020
    from violations 
    join facs using(PWSID)
    where MONITORING_REPORTING = 'Y'
    group by 1,2
  )
  select sum(HIFLD_CAPACITY_2020) as monitoring
  from caps
")
```

## Identify the list of facilities that had a health-based (or, separately, monitoring/reporting) violation in any of the 10 years and sum the capacities of those facilities. Do not double-count facilities for multiple violations. 

```{r}
sqldf("
  with facs as (
    select *
    
    from facilities
    join frs_hifld using (HIFLD_FACILITYID)
    left join pws_reg using (FRS_ID)
  ), caps as (
    select HIFLD_FACILITYID, HIFLD_CAPACITY_2020
    from violations 
    join facs using(PWSID)
    where MONITORING_REPORTING = 'Y'
    or HEALTH_BASED = 'Y'
    or ACUTE_HEALTH_BASED = 'Y'
    group by 1,2
  )
  select sum(HIFLD_CAPACITY_2020) as health_or_monitoring
  from caps
  ")
```